Topological Eulerian Synthesis of Slow Motion Periodic Videos

### 2018-02-02 Paper/paper.tex
Finally wrote up the correct equation for the eigenvalues in my model

Many prior works use a combination of sliding window embeddings and topological data analysis, such as detecting the Lorenz attractor \cite{de2012topological}, quantifying circadian rhythms in gene expression data \cite{perea2015sliding}, detecting chatter in mechanical systems \cite{khasawneh2016chatter}, detecting wheezing in audio \cite{emrani2014real}, quantifying periodic activites in motion capture data \cite{vejdemo2015cohomological, venkataraman2016persistent}, and quantifying vocal fold anomalies in video \cite{tralie2017quasi}, though we believe we are the first to use a sliding window + TDA framework to {\em synthesize} a new result.  Our use of the graph Laplcian to parameterize time series is also unique, though we are inspired by works that have used it to re-arrange images around a loop as a pre-processing step for structure from motion \cite{averbuch2015ringit}, and works which use vector diffusion maps to re-order the frames of microscope images a developing embryo around a loop \cite{dsilva2015diffusionvecordering}.  We use the combination of TDA + Laplacian eigenmaps as a simple alternative to the more complicated cohomology circular coordinates proposed by the authors of \cite{de2011persistent} and used on motion capture data \cite{vejdemo2015cohomological}.

@@

where $\tau$ is a step size and ``dim'' is the dimension of the embedding.  In this work, we fix $\tau = 1$ for simplicity and drop $\tau$ in the subscript, and we also sample times $t$ only at integer frame indices.  As shown by Takens \cite{takens1981detecting}, a sliding window of dimension $2m+1$ of even a single generic observation function (time series) of a dynamical system of intrinsic dimension $m$ is sufficient to reconstruct a topological embedding of the underlying trajectory in the original state space.  For periodic signals, this state space is a torus, and $d$ should be twice the number of harmonics present to injectively reconstruct loops on that torus \cite{perea2015sliding}.  The same is true of videos in general \cite{tralie2017quasi}.  Furthermore, the sliding window length $(\text{dim}-1) \tau$, or simply $\text{dim}-1$ in our case, maximizes the roundness of the embeddings when $\text{dim}-1 = k T$ for some integer $k$, where $T$ is the fundamental period.  Since we will use topological data analysis (TDA) to find the scale of the loop, and since TDA measures roundness, it is important that we tune our window size to be an integer multiple of the period.  To do this, we perform 1D ISOMAP \cite{tenenbaum2000global} on the original frames of the video to generate a 1D ``surrogate signal,'' and we perform an autocorrelation-based fundamental frequency estimation \cite{Mcleod05asmarter} on this signal to estimate the period and, hence, the appropriate dim.  This is quite similar to the diffusion maps + autocorrelation approach in \cite{tralie2017quasi} for analyzing periodic videos.  This estimation often returns integer multiples of the fundamental period, which is fine for our purposes.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{CirculantExample.pdf}
\caption{An example of the unweighted graph Laplacian on an ideal adjacency matrix corresponding to 6 periods of length 14 (top row), and the adjacency matrix from a sliding window embedding of two men doing jumping jacks (bottom row).  The two eigenvectors corresponding the smallest nonzero eigenvalues of the graph Laplacian span sinusoids that are 90 degrees out of phase, and which reconstruct a circle when plotted against each other.}
\label{fig:CirculantExample}
\end{figure}

@@

Now suppose we have performed a sliding window $SW_{\text{dim}}[X[n]]$ for a video $X$, where $X[n] = X[n+kT]$ for some integers $l, k$ so that there are $kT$ points in the sliding window (that is, $X$ repeats itself exactly every $T$ frames $k$ times total), and that $d_{ij} = ||SW_{\text{dim}}[X[i]] - SW_{\text{dim}}[X[n]]||_2$.  In this case, for any $\epsilon > 0$ and $|i - j| = kT$ for some integer $k$, $A_{ij} = 1$.

>>>>

Many prior works use a combination of sliding window embeddings and topological data analysis, such as detecting the Lorenz attractor \cite{de2012topological}, quantifying circadian rhythms in gene expression data \cite{perea2015sliding}, detecting chatter in mechanical systems \cite{khasawneh2016chatter}, detecting wheezing in audio \cite{emrani2014real}, quantifying periodic activites in motion capture data \cite{vejdemo2015cohomological, venkataraman2016persistent}, and quantifying vocal fold anomalies in video \cite{tralie2017quasi}, though we believe we are the first to use a sliding window + TDA framework to {\em synthesize} a new result.  Our use of the graph Laplcian to parameterize time series is also unique, though we are inspired by works that have used it to re-arrange images around a loop as a pre-processing step for structure from motion \cite{averbuch2015ringit}, and works which use vector diffusion maps to re-order the frames of microscope images a developing embryo around a loop \cite{dsilva2015diffusionvecordering}.  One can also view the combination of TDA + Laplacian eigenmaps as a simple alternative to cohomology circular coordinates \cite{de2011persistent,vejdemo2015cohomological} in the special case of periodic data.

@@

where $\tau$ is a step size and ``dim'' is the dimension of the embedding.  In this work, we fix $\tau = 1$ for simplicity and drop $\tau$ in the subscript, and we also sample times $t$ only at integer frame indices.  As shown by Takens \cite{takens1981detecting}, a sliding window of dimension $2m+1$ of even a single generic observation function (time series) of a dynamical system of intrinsic dimension $m$ is sufficient to reconstruct a topological embedding of the underlying trajectory in the original state space.  For periodic signals, this state space is a torus, and $d$ should be twice the number of harmonics present to injectively reconstruct loops on that torus \cite{perea2015sliding}.  The same is true of videos in general \cite{tralie2017quasi}.  Furthermore, the sliding window length $(\text{dim}-1) \tau$, or simply $\text{dim}-1$ in our case, maximizes the roundness of the embeddings when $\text{dim}-1 = k T$ for some integer $k$, where $T$ is the fundamental period.  Since we will use topological data analysis (TDA) to find the scale of the loop, and since TDA measures roundness, it is important that we tune our window size to be an integer multiple of the period.  To do this, we perform 1D ISOMAP \cite{tenenbaum2000global} on the original frames of the video to generate a 1D ``surrogate signal,'' and we perform an autocorrelation-based fundamental frequency estimation \cite{Mcleod05asmarter} on this signal to estimate the period and, hence, the appropriate dim.  This is similar to the diffusion maps + autocorrelation approach in \cite{tralie2017quasi} for analyzing periodic videos.  This estimation often returns integer multiples of the fundamental period, which is fine for our purposes.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{CirculantExample.pdf}
\caption{An example of the unweighted graph Laplacian on an ideal adjacency matrix corresponding to 6 periods of length 14 (top row), and the adjacency matrix from a sliding window embedding of two men doing jumping jacks (bottom row).  }
\label{fig:CirculantExample}
\end{figure}
%The two eigenvectors corresponding the smallest nonzero eigenvalues of the graph Laplacian span sinusoids that are 90 degrees out of phase, and which reconstruct a circle when plotted against each other.

@@

Now suppose we have performed a sliding window $SW_{\text{dim}}[X[n]]$ for a video $X$, where $X[n] = X[n+kT]$ for some integers $l, k$ so that there are $kT$ points in the sliding window (that is, $SW_{\text{dim}}[X[i]]$ repeats itself exactly every $T$ frames $k$ times total, for a total of $N = kT$ windows), and that $d_{ij} = ||SW_{\text{dim}}[X[i]] - SW_{\text{dim}}[X[j]]||_2$.  In this case, for any $\epsilon > 0$ and $|i - j| = kT$ for some integer $k$, $A^u_{\epsilon}(d)_{ij} = 1$.  Let us further suppose it is possible to choose an $\epsilon$ so that each window is connected to the window directly before and after it in time, as well as the windows directly before and after its repetitions, but none others; that is, the unweighted adjacency matrix is a symmetric, {\em circulant matrix}, with the first row $A_{\epsilon}^u(d)_{1j} = 1$ if $j = 1, N-1,$ or $lT, 1 \leq l \leq (k-1)$.  Thus, the Laplacian is also circulant with first row

\begin{equation}
\label{eq:modellaplacian}
L_{1j} = \left\{ \begin{array}{cc} 3k-1 & j = 1 \\ -1 & j = lT, 1 \leq l \leq (k-1) \\ -1 & j = 1, N-1 \\ 0 & \text{otherwise} \end{array} \right\}
\end{equation}

Circulant matrices are diagonalized by the Discrete Fourier Transform \cite{godsil2013algebraic}, and their eigenvalues come in pairs with multiplicity two, with corresopnding eigenvectors $v_1[n] = 1$, $v_{2m}[n] = \cos(2 \pi mn / N)$, $v_{2m+1}[n] = \sin(2 \pi m n / N)$, $m \geq 1$.  In the case of Equation~\ref{eq:modellaplacian}, it can be shown using the DFT that the eigenvalues are $\lambda_1 = 0$ and

\begin{equation}
\begin{array}{cc}\lambda_{2m},\\\lambda_{2m+1}\end{array} = \left\{ \begin{array}{cc} 3k - k\left( 1 + 2 \cos \left( \frac{2 \pi}{kT} m \right) \right) & m = lk, l \in \mathbb{Z}^+ \\ 3k & \text{otherwise}  \end{array} \right\} 
\end{equation}

The smallest two nonzero eigenvalues occur when $m = k$, corresponding to the eigenvectors $v_{2k} = \cos(2 \pi n / T), v_{2k+1} = \sin(2 \pi n / T)$
